---
title: "Project 1"
format: html
editor: visual
---

```{r}
#| label: Import the dataset

here::i_am("Project_1.qmd")

library(tidyverse)
library(rsample)
library(janitor)
library(broom)
library(yardstick)
library(probably)

hospital <- read_csv(here::here("Data/HCAHPS-Hospital.csv"), na = c("Not Applicable", "Not Available"))

```

```{r}
#| label: Filter the dataset for relevent relevent information

hospital <- hospital |>
  clean_names()

hospital_1 <- hospital |>
  select(
    !c("address", "city_town","telephone_number", "hcahps_answer_description") # Not Finished
  )

```

```{r}
#| label: Seperate the data sets

hospital_star_rating <- hospital_1 |> # Star rating is a 1-5 rating of service
  select(
    !c("hcahps_answer_percent", "hcahps_answer_percent_footnote", "hcahps_linear_mean_value" )
  ) |>
  filter(
    !is.na(`patient_survey_star_rating`)
    )


hospital_ans_perc <- hospital_1 |> # Answer percent is percentage of how many times that answer was chosen
  select(
    !c("patient_survey_star_rating", "patient_survey_star_rating_footnote", "hcahps_linear_mean_value" )
  ) |>
  filter(
    !is.na(`hcahps_answer_percent`)
    )


hospital_linear_mean <- hospital_1 |> # linear mean is a standerdization of star rating to be out of 100
  select(
    !c("hcahps_answer_percent", "hcahps_answer_percent_footnote", "patient_survey_star_rating", "patient_survey_star_rating_footnote")
  ) |>
  filter(
    !is.na(`hcahps_linear_mean_value`)
    )
```

```{r}
#| label: Pivot the data

hospital_star_rating_pivot <- hospital_star_rating |>
  pivot_wider(id_cols = `facility_id`,
              names_from = `hcahps_measure_id`,
              values_from = `patient_survey_star_rating`) |>
  clean_names() |>
    rename_with(
    \(x) str_remove(x , "_star_rating")
  )

hospital_ans_perc_pivot <- hospital_ans_perc |>
  pivot_wider(id_cols = `facility_id`,
              names_from = `hcahps_measure_id`,
              values_from = `hcahps_answer_percent`) |>
  clean_names() |>
    rename_with(
    \(x) str_remove(x , "_p")
  )

hospital_linear_mean_pivot <- hospital_linear_mean |>
  pivot_wider(id_cols = `facility_id`,
              names_from = `hcahps_measure_id`,
              values_from = `hcahps_linear_mean_value`) |>
  clean_names() |>
  rename_with(
    \(x) str_remove(x , "_linear_score")
  )

```

```{r}
#| label: Rating vs Recomend analysis

ggplot(
  data = hospital_linear_mean_pivot,
  mapping = aes(x = h_hsp_rating, y = h_recmnd)
) +
  geom_point()

```

There is a positive correlation between the overall hospital rating and the recommendation likelihood. This makes sense since a hospital with a higher rating should be more likely to be recommended. We will be using the overall hospital rating as our response variable. We will judge the other variable to see how much they effect the overall hospital rating.

**DELETE THIS, DO PRINCIPLE COMPONENT ANALYSIS**
